---
title: "Reproducing the advice: WGMIXFISH quality control procedure"
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
- \usepackage[singlelinecheck=false]{caption}
output:
  # rmarkdown::html_vignette:
  html_document:
    number_sections: TRUE
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true

  # rmarkdown::pdf_document:
  #   fig_caption: yes
  #   number_sections: true
  #   df_print: kable
  #   toc: true
  #   toc_depth: 3
  #   keep_tex: yes
vignette: >
  %\VignetteIndexEntry{Reproduce advice}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: refs.bib
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  cache = FALSE, # TRUE for speed. Remember to empty cache after changes to cached R code sections
  cache.path = "cache/", fig.path = "tex/",
  echo = TRUE, 
  collapse = TRUE,
  comment = "#>",
  fig.width = 6, fig.height = 5, dpi=300, 
  out.width="600px", out.height="500px"
)
```

**Required packages to reproduce tutorial:**

```{r packages, message=FALSE, warning=FALSE}
library(FLCore)
library(FLasher)
library(mixfishtools)
library(tidyr)
library(knitr)
```

# Introduction

- Short-term forecast (STF) with deterministic model
- Quality control


# The procedure

1. Start with `FLStock` object with all assessment-estimated values in historical years
2. Extend FLStock into the short-term forecast years using `FLasher::stf` and, if necessary, manually adjust forecast year slot values to match assessment as closely as possible
3. Create `FLasher::fwdControl` object to define the targets and limits of the short-term forecast years
4. Run deterministic forecast with `FLasher::fwd`
5. Compare resulting stock parameters with those reported by the assessment 

An example with saithe (pok.27.3a46)...

## The FLStock with assessment summary

- Contains all assessment estimated values, so that resulting forecast can be reproduced. May differ from the input data in many slots. Here only, catch has been re-estimated to allow for observation error.

```{r FLStockCreation}
data("wgnsskStocks")

stkInp <- wgnsskStocks[["POK_input"]]
stkEst <- wgnsskStocks[["POK"]]

L <- FLStocks(list(input = stkInp, estimated = stkEst))
plot(L) + 
  aes(linetype = stock) +
  scale_color_manual(values = c(8,1)) + 
  scale_linetype_manual(values = c(1,2)) +
  theme_bw()

```

## Extend FLStock for short-term forecast

- extend 3 years (intermediate, advice, and advice +1 years)
- `FLasher::stf` contains several default arguments for defining the number of historical (i.e. terminal) years to average for future values (e.g. mean weights, selection pattern, discard rates). Note that the `wts.nyears` argument defines the averaging period for all biological slots (`catch.wt`, `landings.wt`, `discards.wt`, `stock.wt`, `mat`, `m`, `harvest.spwn`, `m.spwn`).
- If needed, manually adjust slots to reflect the assumptions of the forecast.


```{r FLStockExtention}
yrAssess <- 2023 # final year of assessment data
yrNow <- 2024 # intermediate year
yrTAC <- 2025 # advice year
yrTACp1 <- 2026 # advice year +1 (needed to get SSB at end of yrTAC)

stkProj <- stf(object = stkEst, nyears = 3, wts.nyears = 3, 
  fbar.nyears = 3, f.rescale = TRUE, disc.nyears = 3)

df <- as.data.frame(stkProj)
df <- subset(df, slot %in% c("landings.wt", "discards.wt", "catch.wt", "m", "mat", "harvest") & year > (yrAssess-20))
df$forecast <- df$year %in% c(yrNow, yrTAC, yrTACp1)

ggplot(df) + aes(x = year, y = data, group = age, color = forecast) +
  facet_wrap(~slot, scales = "free_y") +
  geom_line(show.legend = F) + 
  scale_color_manual(values = c(8,1)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) + 
  theme_bw()

```


## Define forecast control

- Set up `fwdControl` to define the targets and limits of the short-term forecast (years)e.g. intermediate year assumptions, target F, etc.)
- Set up stock-recruitment model, which will simply hold the recruitment values used by the forecast. These are typically the reported geometric mean in the advice sheet. 

```{r stfControl}
# stf control (Fsq, followed by 2 years at Fmsy)
Fmsy <- 0.316
ctrl <- fwdControl( 
  data.frame(
    year = c(yrNow, yrTAC, yrTACp1),
    value = c(1, Fmsy, Fmsy),
    quant = c("f"),
    relYear = c(yrAssess, NA, NA)                               
  )
)

# stock-recruitment model (manual input within a geometric mean model)
srPar <- FLPar(c(94047, 94047, 94047), 
  dimnames = list(params="a", year = c(yrNow, yrTAC, yrTACp1), iter = 1))
srMod <- FLSR(model = "geomean", params = srPar)
```

## Run deterministic forecast with FLasher

```{r forecast}
# projection
stkProj <- fwd(object = stkProj, control = ctrl, sr = srMod)

# plot
L <- FLStocks(list(assessment = stkEst, forecast = stkProj[,ac(yrAssess:yrTACp1)]))
plot(L) + 
  # aes(linetype = stock) +
  scale_color_manual(values = c(8,1)) + 
  # scale_linetype_manual(values = c(2,1)) +
  theme_bw()
```



## Compare summary statistics

```{r compare}
# Reported output from single stock headline advice
stfRef <- data.frame(
  model = "NSSK",
  year = 2023:2026,
  catch = c(65865, 78251, 79071, NA),
  landings = c(62691, 74968, 75880, NA),
  fbar = c(0.32, 0.32, 0.316, NA),
  ssb = c(161756, 185632, 195899, 197298)
)

stfDet <- data.frame(
  model = "FLR",
  year = ac(yrAssess:yrTACp1),
  catch = c(catch(stkProj[,ac(yrAssess:yrTACp1)])),
  landings = c(landings(stkProj[,ac(ac(yrAssess:yrTACp1))])),
  fbar = c(fbar(stkProj[,ac(ac(yrAssess:yrTACp1))])),
  ssb = c(ssb(stkProj[,ac(ac(yrAssess:yrTACp1))]))
)  

df <- merge(stfRef, stfDet, all = T)
df <- pivot_longer(df, cols = c(catch, landings, fbar, ssb), 
  names_to = "variable", values_to = "value")
df <- df |>
  filter(
    (variable %in% c("catch", "landings", "fbar") & year <= yrTAC) |
    (variable %in% c("ssb") & year <= yrTACp1))
    

df2 <- pivot_wider(df, names_from = model, values_from = value)
df2$percErr <- round((df2$FLR - df2$NSSK)/df2$NSSK * 100, 1)

ggplot(df) + aes(x = year, y = value, group = model, color = model, shape = model) +
  facet_wrap(~variable, scales = "free_y") +
  geom_line() +
  geom_point(size = 3, stroke = 1) +
  scale_shape_discrete(solid = F) +
  coord_cartesian(ylim = c(0, NA)) +
  theme_bw()


kable(df2, digits = 3)
```


WGMIXISH typically uses a threshold of +/-10%.




# Typical problems

- Forecast slots need additional adjustment to be consistent with the forecast (individual weights, maturity ogive, natural mortality, etc.). State-space models may model future conditions in a more sophisticated way than a simple average of the recent past, such as time-varying processes (e.g. random walk or auto-regressive). These may need to be input manually into the forecast years of the slots. 
- Note that while time-varying selectivity may be largely reproduced in a deterministic forecast, the inclusion of this process in the subsequent mixed fishery model is currently impossible due to the splitting of catches across fleets and métiers.

An example with haddock (had.27.46a20), where additional input for forecast slots was required ...

First, is a forecast withough considering adjustments...

```{r haddockInit}
# load haddock
stkEst <- wgnsskStocks[["HAD"]]

# trim to only historical years
minyear <- range(stkEst)["minyear"]
stkEst <- stkEst[,ac(minyear:yrAssess)]

# simple extension of FLStock
stkProj <- stf(object = stkEst, nyears = 3, wts.nyears = 3, 
  fbar.nyears = 3, f.rescale = TRUE, disc.nyears = 3)

# control
Fmsy <- 0.174
ctrl <- fwdControl( 
  data.frame(
    year = c(yrNow, yrTAC, yrTACp1),
    value = c(1, Fmsy, Fmsy),
    quant = c("f"),
    relYear = c(yrAssess, NA, NA)                               
  )
)

# stock-recruitment model
srPar <- FLPar(c(1788860, 1788860, 1788860), 
  dimnames = list(params="a", year = c(yrNow, yrTAC, yrTACp1), iter = 1))
srMod <- FLSR(model = "geomean", params = srPar)

# projection
stkProj1 <- fwd(object = stkProj, control = ctrl, sr = srMod)

# plot
L <- FLStocks(list(assessment = stkEst, forecast = stkProj1[,ac(yrAssess:yrTACp1)]))
plot(L) + 
  # aes(linetype = stock) +
  scale_color_manual(values = c(8,1)) + 
  # scale_linetype_manual(values = c(2,1)) +
  theme_bw()


# Reported output from single stock headline advice
stfRef <- data.frame(
  model = "NSSK",
  year = 2023:2026,
  catch = c(60979, 66309, 112435, NA),
  landings = c(37702, 53277, 93661, NA),
  fbar = c(0.084, 0.084, 0.174, NA),
  ssb = c(580727, 606378, 535682, 420140)
)

stfDet1 <- data.frame(
  model = "FLR",
  year = ac(yrAssess:yrTACp1),
  catch = c(catch(stkProj1[,ac(yrAssess:yrTACp1)])),
  landings = c(landings(stkProj1[,ac(ac(yrAssess:yrTACp1))])),
  fbar = c(fbar(stkProj1[,ac(ac(yrAssess:yrTACp1))])),
  ssb = c(ssb(stkProj1[,ac(ac(yrAssess:yrTACp1))]))
)  

df1 <- merge(stfRef, stfDet1, all = T)
df1 <- pivot_longer(df1, cols = c(catch, landings, fbar, ssb), 
  names_to = "variable", values_to = "value")
df1 <- df1 |>
  filter(
    (variable %in% c("catch", "landings", "fbar") & year <= yrTAC) |
    (variable %in% c("ssb") & year <= yrTACp1))
df1 <- pivot_wider(df1, names_from = model, values_from = value)
df1$percErr <- round((df1$FLR - df1$NSSK)/df1$NSSK * 100, 1)

ggplot(df1) + aes(x = year, y = percErr) + 
  facet_wrap(~variable) +
  geom_col() +
  geom_hline(yintercept = 0, linetype = 1) +
  geom_hline(yintercept = c(-10,10), linetype = 3) + 
  theme_bw()

```

Then compare to a forecast with furthe adjustments....

```{r haddockFix}
stkProj2 <- stkProj

# replace individual weights
data("haddockWts")
catch.wt(stkProj2)[, ac(yrNow:yrTACp1)][] <- t(haddockWts$catch)
landings.wt(stkProj2)[, ac(yrNow:yrTACp1)][] <- t(haddockWts$landings)
discards.wt(stkProj2)[, ac(yrNow:yrTACp1)][] <- t(haddockWts$discards)
stock.wt(stkProj2)[, ac(yrNow:yrTACp1)][] <- t(haddockWts$stock)

stkProj2 <- fwd(object = stkProj2, control = ctrl, sr = srMod)

stfDet2 <- data.frame(
  model = "FLR",
  year = ac(yrAssess:yrTACp1),
  catch = c(catch(stkProj2[,ac(yrAssess:yrTACp1)])),
  landings = c(landings(stkProj2[,ac(ac(yrAssess:yrTACp1))])),
  fbar = c(fbar(stkProj2[,ac(ac(yrAssess:yrTACp1))])),
  ssb = c(ssb(stkProj2[,ac(ac(yrAssess:yrTACp1))]))
)

df2 <- merge(stfRef, stfDet2, all = T)
df2 <- pivot_longer(df2, cols = c(catch, landings, fbar, ssb), 
  names_to = "variable", values_to = "value")
df2 <- df2 |>
  filter(
    (variable %in% c("catch", "landings", "fbar") & year <= yrTAC) |
    (variable %in% c("ssb") & year <= yrTACp1))
df2 <- pivot_wider(df2, names_from = model, values_from = value)
df2$percErr <- round((df2$FLR - df2$NSSK)/df2$NSSK * 100, 1)
df2$stf <- "adjusted"

df <- df1
df$stf <- "original"
df <- rbind(df, df2)
df$stf <- factor(df$stf, levels = c("original", "adjusted"))

ggplot(df) + aes(x = year, y = percErr, fill = stf) + 
  facet_wrap(~variable) +
  geom_col(position = position_dodge()) +
  geom_hline(yintercept = 0, linetype = 1) +
  geom_hline(yintercept = c(-10,10), linetype = 3) + 
  theme_bw()

```


# A Nephrops example (to come)


# Additional information

## SAM 

For users of SAM, the  `FLfse::SAM2FLStock` function can be used to produce FLStock objects from SAM assessments. By default, this function will return all model inputs as well as non-input fields (e.g. fishing mortality). In order to return all SAM estimated values (e.g. individual weights, catches, maturity, natural mortality), you must specify this with additional arguments (e.g. `SAM2FLStock(object, catch_estimate = T, mat_est = T, stock.wt_est = T, catch.wt_est = T, m_est = T)`)

How to extract information from the forecast object...

```{r samForecast, eval=FALSE, include=FALSE}

Flast <- 0.3204268
Fmsy <- 0.316
TAC_assess_year <- 73815

ARGS <- list(
  fscale = c(NA, NA, NA, NA),
  fval = c(Flast, NA, Fmsy, Fmsy),
  catchval = c(NA, TAC_assess_year, NA, NA),
  fit = pokSAM, ave.years = 2021:2023, rec.years = 2014:2023, 
  nosim = 101,
  label = "TAC (2024), then Fmsy", overwriteSelYears = 2021:2023,
  splitLD = TRUE, savesim = TRUE)

fc <- do.call(stockassessment::forecast, ARGS)
```

